---
title: "Alpha&Beta Diversity"
author: "Daryll Susan Philip"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Load required packages
```{r}
library(readxl)
library(BiocManager)
library(phyloseq)
library(ggplot2)
library(dplyr)
library(vegan)
```

#Alpha Diversity for GC microbes

```{r}
gc_al = read_excel("Gastric Cancer Microbes.xlsx")
gc_al
```

```{r}
alpha_diversity_measures <- function(data, group_col = "Group") {
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }

  data[[group_col]] <- as.factor(data[[group_col]])
  data_no_group <- data[, !(names(data) %in% group_col)]
  data_no_group[] <- lapply(data_no_group, as.numeric)


  if (any(is.na(data_no_group))) {
    stop("The data contains NA values. Please handle them before calculating diversity.")
  }


  if (any(data_no_group < 0)) {
    stop("All abundance values must be non-negative.")
  }

  # Calculate alpha diversity indices
  shannon_diversity <- vegan::diversity(data_no_group, index = "shannon")
  gini_simpson_diversity <- 1 - vegan::diversity(data_no_group, index = "simpson")

  # Create a data frame to store diversity measures
  diversity_measures <- data.frame(
    SampleID = rownames(data),
    Group = data[[group_col]],
    Shannon = shannon_diversity,
    GiniSimpson = gini_simpson_diversity
  )

  groups <- levels(data[[group_col]])
  group_pairs <- combn(groups, 2, simplify = FALSE)

  list(
    diversity_measures = diversity_measures,
    group_pairs = group_pairs
  )
}

results <- alpha_diversity_measures(gc_al, group_col = "Group")
diversity_measures <- results$diversity_measures
group_pairs <- results$group_pairs

print(diversity_measures)
print(group_pairs)

```

#Shannon-Diversity Index for GC

```{r}

# Function to compute Mann-Whitney U test p-values with FDR correction
compute_mann_whitney_pvalues <- function(data, measure, group_col) {
  unique_groups <- unique(data[[group_col]])
  group_comparisons <- combn(unique_groups, 2, simplify = FALSE)
  pvalues <- numeric(length(group_comparisons))
  names(pvalues) <- sapply(group_comparisons, function(x) paste(x, collapse = " vs "))
  
  for (i in seq_along(group_comparisons)) {
    comparison <- group_comparisons[[i]]
    group1_data <- data[data[[group_col]] == comparison[1], measure, drop = TRUE]
    group2_data <- data[data[[group_col]] == comparison[2], measure, drop = TRUE]
    test_result <- wilcox.test(group1_data, group2_data, exact = FALSE)
    pvalues[i] <- test_result$p.value
  }
  
  pvalues <- p.adjust(pvalues, method = "fdr")
  
  return(pvalues)
}


# Function to plot alpha diversity indices without p-value annotations
alpha_diversity_plot <- function(data, measure = "Shannon", group_col = "Group", title = NULL) {
  valid_measures <- c("Shannon", "GiniSimpson")
  if (!(measure %in% valid_measures)) {
    stop("Invalid measure. Choose from: Shannon,GiniSimpson")
  }
  
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }
  
  data[[group_col]] <- as.factor(data[[group_col]])
  if (!(measure %in% colnames(data))) {
    stop(paste("Measure", measure, "is not present in the data."))
  }
  data[[measure]] <- as.numeric(data[[measure]])
  
  num_groups <- length(unique(data[[group_col]]))
  palette <- if (num_groups == 2) c("#D95F02", "#1B9E77") else c("#3288bd", "#d53e4f", "#62954C", "#C59144")
  
  plot <- ggpubr::ggviolin(
    data,
    x = group_col,
    y = measure,
    color = "black",
    add = "boxplot",
    fill = group_col,
    palette = palette,
    legend = "none"
  ) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 1) +
    ggplot2::xlab("") +
    ggplot2::ylab(toupper(measure)) +
    ggthemes::theme_pander() +
    ggplot2::theme(
      panel.grid.major = ggplot2::element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor = ggplot2::element_line(color = "grey90", linewidth = 0.25),
      axis.text.x = element_text(size = 18),
      axis.text.y = element_text(size = 18),
      axis.title.y = element_text(size = 18, angle = 90, margin = margin(l = 10))
    )
  
  if (!is.null(title)) {
    plot <- plot + ggplot2::ggtitle(title)
  }
  
  return(plot)
}

# Compute p-values separately
p_values <- compute_mann_whitney_pvalues(data = diversity_measures, measure = "Shannon", group_col = "Group")
print(p_values)  # Print the p-values

# Plot the graph without p-value annotations
plot <- alpha_diversity_plot(data = diversity_measures, measure = "Shannon", group_col = "Group", title = "Shannon Diversity Index GC")
print(plot)

```

#Gini-Simpson Index for GC

```{r}


# Function to plot alpha diversity indices without p-value annotations 
alpha_diversity_plot <- function(data, measure = "GiniSimpson", group_col = "Group", title = NULL) {
  valid_measures <- c("Shannon","GiniSimpson")
  if (!(measure %in% valid_measures)) {
    stop("Invalid measure. Choose from: Shannon, GiniSimpson")
  }
  
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }
  
  data[[group_col]] <- as.factor(data[[group_col]])
  if (!(measure %in% colnames(data))) {
    stop(paste("Measure", measure, "is not present in the data."))
  }
  data[[measure]] <- as.numeric(data[[measure]])
  
  num_groups <- length(unique(data[[group_col]]))
  palette <- if (num_groups == 2) c("#D95F02", "#1B9E77") else c("#3288bd", "#d53e4f", "#62954C", "#C59144")
  
  plot <- ggpubr::ggviolin(
    data,
    x = group_col,
    y = measure,
    color = "black",
    add = "boxplot",
    fill = group_col,
    palette = palette,
    legend = "none"
  ) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 1) +
    ggplot2::xlab("") +
    ggplot2::ylab(toupper(measure)) +
    ggthemes::theme_pander() +
    ggplot2::theme(
      panel.grid.major = ggplot2::element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor = ggplot2::element_line(color = "grey90", linewidth = 0.25),
      axis.text.x = element_text(size = 18),
      axis.text.y = element_text(size = 18),
      axis.title.y = element_text(size = 18, angle = 90, margin = margin(l = 10))
    )
  
  if (!is.null(title)) {
    plot <- plot + ggplot2::ggtitle(title)
  }
  
  return(plot)
}

# Compute p-values separately
p_values <- compute_mann_whitney_pvalues(data = diversity_measures, measure = "GiniSimpson", group_col = "Group")
print(p_values)  

# Plot the graph without p-value annotations
plot <- alpha_diversity_plot(data = diversity_measures, measure = "GiniSimpson", group_col = "Group", title = "Gini Simpson Diversity Index GC")
print(plot)

```

#Beta Diversity for GC

```{r}

ge_datas <- gc_al %>%
  select(-Group) %>%
  mutate_if(is.character, as.numeric)

ge_jaccard <- vegdist(ge_datas, method = "jaccard") # Calculate the Jaccard index
ge_nmds <- metaMDS(ge_jaccard, distance = "jaccard", k = 4, trymax = 1000) # Perform NMDS
stress_value <- ge_nmds$stress


plot_df <- as.data.frame(scores(ge_nmds, display = "sites"))
plot_df$Group <- gc_alp$Group


pal <- c("Healthy" = "#1B9E77", "Gastrectomy" = "#D95F02")

# Create the NMDS plot
plot_nmds <- ggplot(plot_df, aes(x = NMDS1, y = NMDS2, color = Group, shape = Group)) +
  geom_point(size = 4, alpha = 1, , shape = 16) +
  scale_color_manual(values = pal) +
  stat_ellipse(linetype = 2, linewidth = 1.3) +
  theme_minimal() +
  labs(
    title = "NMDS of Jaccard Distances for Gastric Cancer",
    x = "NMDS1",
    y = "NMDS2"
  ) +
  theme(
    axis.title.x = element_text(size = 16),  # Increase x-axis title size
    axis.title.y = element_text(size = 16),  # Increase y-axis title size
    axis.text = element_text(size = 14),     # Increase axis text size
    legend.text = element_text(size = 14),    # Increase legend text size
    legend.title = element_text(size = 16)    # Increase legend title size
  )   

print(plot_nmds)
print(stress_value)

```

#Alpha Diversity for CRC microbes

```{r}
cc_al = read_excel("CC-gene.xlsx")
cc_al
```

```{r}
alpha_diversity_measures <- function(data, group_col = "Group") {
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }

  data[[group_col]] <- as.factor(data[[group_col]])
  data_no_group <- data[, !(names(data) %in% group_col)]
  data_no_group[] <- lapply(data_no_group, as.numeric)


  if (any(is.na(data_no_group))) {
    stop("The data contains NA values. Please handle them before calculating diversity.")
  }


  if (any(data_no_group < 0)) {
    stop("All abundance values must be non-negative.")
  }

  # Calculate alpha diversity indices
  shannon_diversity <- vegan::diversity(data_no_group, index = "shannon")
  gini_simpson_diversity <- 1 - vegan::diversity(data_no_group, index = "simpson")

  # Create a data frame to store diversity measures
  diversity_measures <- data.frame(
    SampleID = rownames(data),
    Group = data[[group_col]],
    Shannon = shannon_diversity,
    GiniSimpson = gini_simpson_diversity
  )

  groups <- levels(data[[group_col]])
  group_pairs <- combn(groups, 2, simplify = FALSE)

  list(
    diversity_measures = diversity_measures,
    group_pairs = group_pairs
  )
}

results <- alpha_diversity_measures(cc_al, group_col = "Group")
diversity_measures <- results$diversity_measures
group_pairs <- results$group_pairs

print(diversity_measures)
print(group_pairs)

```


#Shannon-Diversity Index for IBD

```{r}
# Function to compute Mann-Whitney U test p-values with FDR correction
compute_mann_whitney_pvalues <- function(data, measure, group_col) {
  unique_groups <- unique(data[[group_col]])
  group_comparisons <- combn(unique_groups, 2, simplify = FALSE)
  pvalues <- numeric(length(group_comparisons))
  names(pvalues) <- sapply(group_comparisons, function(x) paste(x, collapse = " vs "))
  
  for (i in seq_along(group_comparisons)) {
    comparison <- group_comparisons[[i]]
    group1_data <- data[data[[group_col]] == comparison[1], measure, drop = TRUE]
    group2_data <- data[data[[group_col]] == comparison[2], measure, drop = TRUE]
    test_result <- wilcox.test(group1_data, group2_data, exact = FALSE)
    pvalues[i] <- test_result$p.value
  }
  
  pvalues <- p.adjust(pvalues, method = "fdr")
  
  return(pvalues)
}


# Function to plot alpha diversity indices without p-value annotations
alpha_diversity_plot <- function(data, measure = "Shannon", group_col = "Group", title = NULL) {
  valid_measures <- c("Shannon", "GiniSimpson")
  if (!(measure %in% valid_measures)) {
    stop("Invalid measure. Choose from: Shannon,GiniSimpson")
  }
  
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }
  
  data[[group_col]] <- as.factor(data[[group_col]])
  if (!(measure %in% colnames(data))) {
    stop(paste("Measure", measure, "is not present in the data."))
  }
  data[[measure]] <- as.numeric(data[[measure]])
  
  num_groups <- length(unique(data[[group_col]]))
  palette <- if (num_groups == 2) c("#D95F02", "#1B9E77") else c("#3288bd", "#d53e4f", "#62954C", "#C59144")
  
  plot <- ggpubr::ggviolin(
    data,
    x = group_col,
    y = measure,
    color = "black",
    add = "boxplot",
    fill = group_col,
    palette = palette,
    legend = "none"
  ) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 1) +
    ggplot2::xlab("") +
    ggplot2::ylab(toupper(measure)) +
    ggthemes::theme_pander() +
    ggplot2::theme(
      panel.grid.major = ggplot2::element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor = ggplot2::element_line(color = "grey90", linewidth = 0.25),
      axis.text.x = element_text(size = 18),
      axis.text.y = element_text(size = 18),
      axis.title.y = element_text(size = 18, angle = 90, margin = margin(l = 10))
    )
  
  if (!is.null(title)) {
    plot <- plot + ggplot2::ggtitle(title)
  }
  
  return(plot)
}

# Compute p-values separately
p_values <- compute_mann_whitney_pvalues(data = diversity_measures, measure = "Shannon", group_col = "Group")
print(p_values)  

# Plot the graph without p-value annotations
plot <- alpha_diversity_plot(data = diversity_measures, measure = "Shannon", group_col = "Group", title = "Shannon Diversity Index CRC")
print(plot)

```

#Gini-Simpson Index for CRC

```{r}


# Function to plot alpha diversity indices without p-value annotations 
alpha_diversity_plot <- function(data, measure = "GiniSimpson", group_col = "Group", title = NULL) {
  valid_measures <- c("Shannon","GiniSimpson")
  if (!(measure %in% valid_measures)) {
    stop("Invalid measure. Choose from: Shannon, GiniSimpson")
  }
  
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }
  
  data[[group_col]] <- as.factor(data[[group_col]])
  if (!(measure %in% colnames(data))) {
    stop(paste("Measure", measure, "is not present in the data."))
  }
  data[[measure]] <- as.numeric(data[[measure]])
  
  num_groups <- length(unique(data[[group_col]]))
  palette <- if (num_groups == 2) c("#D95F02", "#1B9E77") else c("#3288bd", "#d53e4f", "#62954C", "#C59144")
  
  plot <- ggpubr::ggviolin(
    data,
    x = group_col,
    y = measure,
    color = "black",
    add = "boxplot",
    fill = group_col,
    palette = palette,
    legend = "none"
  ) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 1) +
    ggplot2::xlab("") +
    ggplot2::ylab(toupper(measure)) +
    ggthemes::theme_pander() +
    ggplot2::theme(
      panel.grid.major = ggplot2::element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor = ggplot2::element_line(color = "grey90", linewidth = 0.25),
      axis.text.x = element_text(size = 18),
      axis.text.y = element_text(size = 18),
      axis.title.y = element_text(size = 18, angle = 90, margin = margin(l = 10))
    )
  
  if (!is.null(title)) {
    plot <- plot + ggplot2::ggtitle(title)
  }
  
  return(plot)
}


p_values <- compute_mann_whitney_pvalues(data = diversity_measures, measure = "GiniSimpson", group_col = "Group")
print(p_values)  

# Plot the graph without p-value annotations
plot <- alpha_diversity_plot(data = diversity_measures, measure = "GiniSimpson", group_col = "Group", title = "Gini Simpson Diversity Index CRC")
print(plot)

```


#Beta Diversity for CRC

```{r}

ge_datas <- cc_al %>%
  select(-Group) %>%
  mutate_if(is.character, as.numeric)


ge_jaccard <- vegdist(ge_datas, method = "jaccard") # Calculate the Jaccard index
ge_nmds <- metaMDS(ge_jaccard, distance = "jaccard", k = 4, trymax = 1000) # Perform NMDS
stress_value <- ge_nmds$stress
plot_df <- as.data.frame(scores(ge_nmds, display = "sites"))
plot_df$Group <- cc_al$Group
pal <- c("Healthy" = "#1B9E77", "Colon Cancer" = "#D95F02")

#Plot NMDS
plot_nmds <- ggplot(plot_df, aes(x = NMDS1, y = NMDS2, color = Group, shape = Group)) +
  geom_point(size = 4, alpha = 1, shape = 16) +
  scale_color_manual(values = pal) +
  stat_ellipse(linetype = 2, linewidth = 1.3) + 
  theme_minimal() +
  labs(
    title = "NMDS of Jaccard Distances for Colon Cancer",
    x = "NMDS1",
    y = "NMDS2"
  ) +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  )


print(plot_nmds)
print(stress_value)

```



#Alpha Diversity for IBD microbes

```{r}
ibd_al = read_excel("IBD-gene.xlsx")
ibd_al
```

```{r}
alpha_diversity_measures <- function(data, group_col = "Group") {
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }

  data[[group_col]] <- as.factor(data[[group_col]])
  data_no_group <- data[, !(names(data) %in% group_col)]
  data_no_group[] <- lapply(data_no_group, as.numeric)


  if (any(is.na(data_no_group))) {
    stop("The data contains NA values. Please handle them before calculating diversity.")
  }


  if (any(data_no_group < 0)) {
    stop("All abundance values must be non-negative.")
  }

  # Calculate alpha diversity indices
  shannon_diversity <- vegan::diversity(data_no_group, index = "shannon")
  gini_simpson_diversity <- 1 - vegan::diversity(data_no_group, index = "simpson")

  # Create a data frame to store diversity measures
  diversity_measures <- data.frame(
    SampleID = rownames(data),
    Group = data[[group_col]],
    Shannon = shannon_diversity,
    GiniSimpson = gini_simpson_diversity
  )

  groups <- levels(data[[group_col]])
  group_pairs <- combn(groups, 2, simplify = FALSE)

  list(
    diversity_measures = diversity_measures,
    group_pairs = group_pairs
  )
}

results <- alpha_diversity_measures(ibd_al, group_col = "Group")
diversity_measures <- results$diversity_measures
group_pairs <- results$group_pairs

print(diversity_measures)
print(group_pairs)

```


#Shannon-Diversity Index for IBD

```{r}
# Function to compute Mann-Whitney U test p-values with FDR correction
compute_mann_whitney_pvalues <- function(data, measure, group_col) {
  unique_groups <- unique(data[[group_col]])
  group_comparisons <- combn(unique_groups, 2, simplify = FALSE)
  pvalues <- numeric(length(group_comparisons))
  names(pvalues) <- sapply(group_comparisons, function(x) paste(x, collapse = " vs "))
  
  for (i in seq_along(group_comparisons)) {
    comparison <- group_comparisons[[i]]
    group1_data <- data[data[[group_col]] == comparison[1], measure, drop = TRUE]
    group2_data <- data[data[[group_col]] == comparison[2], measure, drop = TRUE]
    test_result <- wilcox.test(group1_data, group2_data, exact = FALSE)
    pvalues[i] <- test_result$p.value
  }
  
  pvalues <- p.adjust(pvalues, method = "fdr")
  
  return(pvalues)
}


# Function to plot alpha diversity indices without p-value annotations
alpha_diversity_plot <- function(data, measure = "Shannon", group_col = "Group", title = NULL) {
  valid_measures <- c("Shannon", "GiniSimpson")
  if (!(measure %in% valid_measures)) {
    stop("Invalid measure. Choose from: Shannon,GiniSimpson")
  }
  
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }
  
  data[[group_col]] <- as.factor(data[[group_col]])
  if (!(measure %in% colnames(data))) {
    stop(paste("Measure", measure, "is not present in the data."))
  }
  data[[measure]] <- as.numeric(data[[measure]])
  
  num_groups <- length(unique(data[[group_col]]))
  palette <- if (num_groups == 2) c("#D95F02", "#1B9E77") else c("#3288bd", "#d53e4f", "#62954C", "#C59144")
  
  plot <- ggpubr::ggviolin(
    data,
    x = group_col,
    y = measure,
    color = "black",
    add = "boxplot",
    fill = group_col,
    palette = palette,
    legend = "none"
  ) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 1) +
    ggplot2::xlab("") +
    ggplot2::ylab(toupper(measure)) +
    ggthemes::theme_pander() +
    ggplot2::theme(
      panel.grid.major = ggplot2::element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor = ggplot2::element_line(color = "grey90", linewidth = 0.25),
      axis.text.x = element_text(size = 18),
      axis.text.y = element_text(size = 18),
      axis.title.y = element_text(size = 18, angle = 90, margin = margin(l = 10))
    )
  
  if (!is.null(title)) {
    plot <- plot + ggplot2::ggtitle(title)
  }
  
  return(plot)
}

# Compute p-values separately
p_values <- compute_mann_whitney_pvalues(data = diversity_measures, measure = "Shannon", group_col = "Group")
print(p_values)  

# Plot the graph without p-value annotations
plot <- alpha_diversity_plot(data = diversity_measures, measure = "Shannon", group_col = "Group", title = "Shannon Diversity Index IBD")
print(plot)

```

#Gini-Simpson Index for IBD

```{r}


# Function to plot alpha diversity indices without p-value annotations 
alpha_diversity_plot <- function(data, measure = "GiniSimpson", group_col = "Group", title = NULL) {
  valid_measures <- c("Shannon","GiniSimpson")
  if (!(measure %in% valid_measures)) {
    stop("Invalid measure. Choose from: Shannon, GiniSimpson")
  }
  
  if (!group_col %in% colnames(data)) {
    stop("The specified group column does not exist in the data.")
  }
  
  data[[group_col]] <- as.factor(data[[group_col]])
  if (!(measure %in% colnames(data))) {
    stop(paste("Measure", measure, "is not present in the data."))
  }
  data[[measure]] <- as.numeric(data[[measure]])
  
  num_groups <- length(unique(data[[group_col]]))
  palette <- if (num_groups == 2) c("#D95F02", "#1B9E77") else c("#3288bd", "#d53e4f", "#62954C", "#C59144")
  
  plot <- ggpubr::ggviolin(
    data,
    x = group_col,
    y = measure,
    color = "black",
    add = "boxplot",
    fill = group_col,
    palette = palette,
    legend = "none"
  ) +
    geom_jitter(shape = 16, position = position_jitter(0.2), size = 1) +
    ggplot2::xlab("") +
    ggplot2::ylab(toupper(measure)) +
    ggthemes::theme_pander() +
    ggplot2::theme(
      panel.grid.major = ggplot2::element_line(color = "grey80", linewidth = 0.5),
      panel.grid.minor = ggplot2::element_line(color = "grey90", linewidth = 0.25),
      axis.text.x = element_text(size = 18),
      axis.text.y = element_text(size = 18),
      axis.title.y = element_text(size = 18, angle = 90, margin = margin(l = 10))
    )
  
  if (!is.null(title)) {
    plot <- plot + ggplot2::ggtitle(title)
  }
  
  return(plot)
}


p_values <- compute_mann_whitney_pvalues(data = diversity_measures, measure = "GiniSimpson", group_col = "Group")
print(p_values)  

# Plot the graph without p-value annotations
plot <- alpha_diversity_plot(data = diversity_measures, measure = "GiniSimpson", group_col = "Group", title = "Gini Simpson Diversity Index IBD")
print(plot)

```

#Beta Diversity in IBD

```{r}

ge_datas <- ibd_al %>%
  select(-Group) %>%
  mutate_if(is.character, as.numeric)
ge_jaccard <- vegdist(ge_datas, method = "jaccard") # Calculate the Jaccard index
ge_nmds <- metaMDS(ge_jaccard, distance = "jaccard", k = 4, trymax = 1000) # Perform NMDS
stress_value <- ge_nmds$stress


plot_df <- as.data.frame(scores(ge_nmds, display = "sites"))
plot_df$Group <- ibd_al$Group
pal <- c("IBD" = "#D95F02","Healthy" = "#1B9E77")

plot_nmds <- ggplot(plot_df, aes(x = NMDS1, y = NMDS2, color = Group, shape = Group)) +
  geom_point(size = 4, alpha = 1, shape = 16) +
  scale_color_manual(values = pal) +
  stat_ellipse(linetype = 2, linewidth = 1.3) +  
  theme_minimal() +
  labs(
    title = "NMDS of Jaccard Distances for IBD",
    x = "NMDS1",
    y = "NMDS2"
  ) +
  theme(
    axis.title.x = element_text(size = 16),
    axis.title.y = element_text(size = 16),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 14),
    legend.title = element_text(size = 16)
  ) 


# Print the plot
print(plot_nmds)
print(stress_value)

```

