---
title: "Colon Cancer Preprocessing"
author: "Daryll Susan Philip"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/DaryllPhilip/Desktop/THESIS /Colon Cancer/YACHIDA_CRC_2019")
```

```{r}
getwd()
```


#Load required packages
```{r}
library(tidyverse)
library(dplyr)
library(vegan)
library(tidyr)
library(naniar)
library(ape)
library(ggplot2)
library(FactoMineR)
library(factoextra)
library(stats)
```


#Load the datasets
```{r}
#microbiome
gene_cc = read.csv("gene_sorted.csv", header = TRUE, stringsAsFactors = FALSE)
gene_cc

#metabolome
mtb_cc = read.csv("mtb_sorted.csv", header = TRUE, stringsAsFactors = FALSE)
mtb_cc
```


#Microbiome CRC

```{r}

# Count the number of patients in each group
patient_counts <- gene_cc %>%
  group_by(Study.Group) %>%
  summarise(Count = n())
print(patient_counts)


colon_cancer_count <- sum(gene_cc$Study.Group == "Colon Cancer")
healthy_count <- sum(gene_cc$Study.Group == "Healthy")

cat("Number of Colon Cancer patients:", colon_cancer_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

```


#Check for missing values
```{r}

missing_summary <- gene_cc %>%
  summarise_all(~ sum(is.na(.))) %>%
  gather(key = "Column", value = "MissingValues") %>%
  mutate(MissingPercentage = (MissingValues / nrow(gene_cc)) * 100)
missing_summary <- naniar::miss_var_summary(gene_cc)

print(missing_summary)
naniar::gg_miss_var(gene_cc)

#there were no missing values
```

#Identify sparse features
```{r}


sparsity <- colSums(gene_cc == 0) / nrow(gene_cc) * 100
bins <- cut(sparsity, breaks = seq(0, 110, by = 10), include.lowest = TRUE, right = FALSE, labels = FALSE)
sparsity_counts <- table(bins)
bin_labels <- c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")
total_features <- length(sparsity)
surviving_features <- sum(bins == 1) 

# Print the table
print("Sparsity (%) | Number of Features")
for (i in seq_along(bin_labels)) {
  count <- ifelse(i %in% names(sparsity_counts), sparsity_counts[as.character(i)], 0)
  cat(bin_labels[i], "|", count, "\n")
}


cat("\nTotal number of microbiome features:", total_features, "\n")
cat("Number of surviving microbiome features (non-sparse):", surviving_features, "\n")
```

#Removing sparse features
```{r}

sparsity <- colSums(gene_cc == 0) / nrow(gene_cc) * 100
gene_new <- gene_cc[, sparsity <= 70]

cat("Original number of features:", ncol(gene_cc), "\n")
cat("Number of features after filtering (<= 70% sparsity):", ncol(gene_new), "\n")

```

#Min-max scale the features
```{r}
min_max_scaling <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

gene_new <- gene_new %>%
  mutate(across(where(is.numeric), min_max_scaling))
gene_new
```


#PCoA to visualise the outliers

```{r}

gene_new <- gene_new[, !(names(gene_new) %in% c("Study.Group"))]
bray_curtis_dist <- vegdist(gene_new, method = "bray")
pcoa_results <- pcoa(bray_curtis_dist)
coordinates <- pcoa_results$vectors[, 1:2]
colnames(coordinates) <- c("Axis.1", "Axis.2")
pcoa_df <- as.data.frame(coordinates)
pcoa_df$Study.Group <- gene_new$Study.Group 
pcoa_df$Study.Group <- as.factor(pcoa_df$Study.Group)


pcoa_df <- pcoa_df %>%
  mutate(OutlierStatus = ifelse(Outlier, "Outlier", "Non-Outlier"))
print(table(pcoa_df$OutlierStatus))

#Plot PCoA with outliers highlighted
ggplot(pcoa_df, aes(x = Axis.1, y = Axis.2, color = Study.Group)) +
  geom_point(aes(shape = OutlierStatus), size = 3) +  # Shape based on outlier status
  stat_ellipse(aes(group = Study.Group, color = Study.Group), type = "t", level = 0.95) +
  scale_color_manual(values = c("Colon Cancer" = "#D95F02", "Healthy" = "#1B9E77")) +
  labs(title = "PCoA For Genera (Bray-Curtis) with Outliers",
       x = paste("PCoA1 (", round(100 * pcoa_results$values$Relative_eig[1], 2), "%)", sep = ""),
       y = paste("PCoA2 (", round(100 * pcoa_results$values$Relative_eig[2], 2), "%)", sep = ""),
       color = "Group", shape = "Outlier Status") +
  theme_minimal()

```

#Remove the outliers and replot the graph 
```{r}

non_outlier_indices <- which(pcoa_df$OutlierStatus == "Non-Outlier")
cleaned_gene <- gene_new[non_outlier_indices, ]
gen_cleaned <- cleaned_gene[, !(names(cleaned_gene) %in% c("Study.Group"))]
bray_curtis_dist_cleaned <- vegdist(gen_cleaned, method = "bray")

# Perform PCoA on the cleaned Bray-Curtis dissimilarity matrix
pcoa_results_cleaned <- pcoa(bray_curtis_dist_cleaned)
coordinates_cleaned <- pcoa_results_cleaned$vectors[, 1:2]
colnames(coordinates_cleaned) <- c("Axis.1", "Axis.2")
pcoa_df_cleaned <- as.data.frame(coordinates_cleaned)
pcoa_df_cleaned$Study.Group <- cleaned_gene$Study.Group  # Ensure this column exists in 'cleaned_gene'
pcoa_df_cleaned$Study.Group <- as.factor(pcoa_df_cleaned$Study.Group)

#Plot PCoA for cleaned dataset with outliers removed
ggplot(pcoa_df_cleaned, aes(x = Axis.1, y = Axis.2, color = Study.Group )) +
  geom_point(size = 3) +  
  stat_ellipse(aes(group = Study.Group, color = Study.Group), type = "t", level = 0.95) +
  scale_color_manual(values = c("Colon Cancer" = "#D95F02", "Healthy" = "#1B9E77")) +
  labs(title = "PCoA For Colon Cancer Genera (Bray-Curtis)",
       x = paste("PCoA1 (", round(100 * pcoa_results_cleaned$values$Relative_eig[1], 2), "%)", sep = ""),
       y = paste("PCoA2 (", round(100 * pcoa_results_cleaned$values$Relative_eig[2], 2), "%)", sep = ""),
       color = "Group") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 14),  
    axis.title.y = element_text(size = 14), 
    legend.title = element_text(size = 12)  
  )

```

#Check for number of colon cancer and healthy patients for microbiome dataset
```{r}
gene_pcoa = cleaned_gene
patient_counts <- gene_pcoa %>%
  group_by(Study.Group) %>%
  summarise(Count = n())
print(patient_counts)


cc_count <- sum(gene_pcoa$Study.Group == "Colon Cancer")
healthy_count <- sum(gene_pcoa$Study.Group == "Healthy")

cat("Number of Colon Cancer patients:", cc_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

# 267 samples from the microbiome dataset were considered for final analysis from the initial 277 samples
```


#Kruskal-Wallis test for Genera/Microbiome

```{r}

cc_ge <- gene_pcoa[, !colnames(gene_pcoa) %in% "Study.Group"]
cc_ge$Study.Group <- as.factor(cc_gene$Study.Group)
long_data <- cc_gene %>%
  pivot_longer(cols = -Study.Group, names_to = "Compound", values_to = "Value")

# Perform Kruskal-Wallis test for each compound
compounds <- unique(long_data$Compound)
kruskal_results <- lapply(compounds, function(compound) {
  subset_data <- long_data %>% filter(Compound == compound)
  
  if (length(unique(subset_data$Study.Group)) < 2) {
    warning(paste("Skipping compound", compound, "due to insufficient group data"))
    return(NULL)
  }
  
  result <- kruskal.test(Value ~ Study.Group, data = subset_data)
  return(list(Compound = compound, Kruskal_Result = result))
})

results_gene <- data.frame(
  Compound = character(),
  H_Statistic = numeric(),
  Degrees_Freedom = integer(),
  P_Value = numeric(),
  P_Value_FDR = numeric(),  
  Reject_Null = logical(),  
  stringsAsFactors = FALSE
)

for (res in kruskal_results) {
  if (!is.null(res)) {
    p_val <- res$Kruskal_Result$p.value
    p_val_fdr <- p.adjust(p_val, method = "fdr") # FDR corrected p-value
    reject_null <- p_val_fdr < 0.05
    
    results_gene <- rbind(results_gene, data.frame(
      Compound = res$Compound,
      H_Statistic = res$Kruskal_Result$statistic,
      Degrees_Freedom = res$Kruskal_Result$parameter,
      P_Value = p_val,
      P_Value_FDR = p_val_fdr,
      Reject_Null = reject_null
    ))
  }
}

num_rejected <- sum(results_gene$Reject_Null)
num_accepted <- nrow(results_gene) - num_rejected

cat("Number of null hypotheses rejected:", num_rejected, "\n")
cat("Number of null hypotheses accepted:", num_accepted, "\n")
```

# Remove insignificant features
```{r}
non_significant_compounds <- results_gene$Compound[results_gene$Reject_Null == FALSE]
gene_filtered <- gene_pcoa[, !names(gene_pcoa) %in% non_significant_compounds]
print(gene_filtered)
#the whole process left us with 206 microbes for machine learning analysis
```


#Metabolites(CC)

```{r}

# Count the number of patients in each group
patient_counts <- mtb_cc %>%
  group_by(Study.Group) %>%
  summarise(Count = n())
print(patient_counts)


colon_cancer_count <- sum(mtb_cc$Study.Group == "Colon Cancer")
healthy_count <- sum(mtb_cc$Study.Group == "Healthy")

cat("Number of Colon Cancer patients:", colon_cancer_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

```


#Check for missing values
```{r}

missing_summary <- mtb_cc %>%
  summarise_all(~ sum(is.na(.))) %>%
  gather(key = "Column", value = "MissingValues") %>%
  mutate(MissingPercentage = (MissingValues / nrow(mtb_cc)) * 100)
missing_summary <- naniar::miss_var_summary(mtb_cc)

print(missing_summary)
naniar::gg_miss_var(mtb_cc)

#there were no missing values
```

#Identify sparse features
```{r}


sparsity <- colSums(mtb_cc == 0) / nrow(mtb_cc) * 100
bins <- cut(sparsity, breaks = seq(0, 110, by = 10), include.lowest = TRUE, right = FALSE, labels = FALSE)
sparsity_counts <- table(bins)
bin_labels <- c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")
total_features <- length(sparsity)
surviving_features <- sum(bins == 1) 

# Print the table
print("Sparsity (%) | Number of Features")
for (i in seq_along(bin_labels)) {
  count <- ifelse(i %in% names(sparsity_counts), sparsity_counts[as.character(i)], 0)
  cat(bin_labels[i], "|", count, "\n")
}


cat("\nTotal number of microbiome features:", total_features, "\n")
cat("Number of surviving microbiome features (non-sparse):", surviving_features, "\n")
```

#Removing sparse features
```{r}

sparsity <- colSums(mtb_cc == 0) / nrow(mtb_cc) * 100
mtb_new <- mtb_cc[, sparsity <= 80]

cat("Original number of features:", ncol(mtb_cc), "\n")
cat("Number of features after filtering (<= 80% sparsity):", ncol(mtb_new), "\n")

```

#Min-max scale the features
```{r}
min_max_scaling <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

mtb_new <- mtb_new %>%
  mutate(across(where(is.numeric), min_max_scaling))
mtb_new
```

#PCA for metabolites

```{r}


mtbs <- mtb_new[, !(names(mtb_new) %in% c("Study.Group"))]


mtbs <- mtbs[, apply(mtbs, 2, function(col) sd(col) != 0)]
pca_result_mtb <- prcomp(mtbs, scale. = TRUE)
fviz_eig(pca_result_mtb, addlabels = TRUE, ylim = c(0, 70))
explained_variance <- summary(pca_result_mtb)$importance[2, ]
pca_df <- as.data.frame(pca_result_mtb$x)



#Calculate Mahalanobis distance
mean_pca <- colMeans(pca_df[, 1:2])
cov_pca <- cov(pca_df[, 1:2])
mahalanobis_dist <- apply(pca_df[, 1:2], 1, function(x) mahalanobis(x, mean_pca, cov_pca))
cutoff <- qchisq(0.95, df = 2)
pca_df$Outlier <- mahalanobis_dist > cutoff


pca_df <- pca_df %>%
  mutate(OutlierStatus = ifelse(Outlier, "Outlier", "Non-Outlier"))

#Plot PCA 
ggplot(pca_df, aes(x = PC1, y = PC2, color = Study.Group, shape = OutlierStatus)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Non-Outlier" = 16, "Outlier" = 4)) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "PCA For Metabolites",
       x = paste("PC1 (", round(100 * explained_variance[1], 2), "%)", sep = ""),
       y = paste("PC2 (", round(100 * explained_variance[2], 2), "%)", sep = ""),
       color = "Study.Group",
       shape = "Outlier Status") +
  theme_minimal()

table(pca_df$OutlierStatus)

```

#Remove outliers and replot the graph

```{r}

pca_df_cleaned <- pca_df %>%
  filter(OutlierStatus == "Non-Outlier")

mtb_cc <- mtb_new %>%
  filter(row.names(mtb_new) %in% row.names(pca_df_cleaned))

#Plot PCA with outliers removed
ggplot(pca_df_cleaned, aes(x = PC1, y = PC2, color = Study.Group)) +
  geom_point(size = 3) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "PCA For Colon Cancer Metabolites ",
       x = paste("PC1 (", round(100 * explained_variance[1], 2), "%)", sep = ""),
       y = paste("PC2 (", round(100 * explained_variance[2], 2), "%)", sep = ""),
       color = "Group") +
  theme_minimal() +
theme(
    plot.title = element_text(size = 16),  
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    legend.title = element_text(size = 12)  
  )
```

#Check for number of Colon Cancer and Healthy patients for metabolome dataset
```{r}
mtb_pcoa = mtb_cc
patient_counts <- mtb_pcoa %>%
  group_by(Study.Group) %>%
  summarise(Count = n())
print(patient_counts)


cc_count <- sum(mtb_pcoa$Study.Group == "Colon Cancer")
healthy_count <- sum(mtb_pcoa$Study.Group == "Healthy")

cat("Number of Colon Cancer patients:", cc_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

# 265 samples from the metabolome dataset were considered for final analysis from the previous 277 samples
```

#Kruskal-Wallis test for Metabolome

```{r}

cc_mtb <- mtb_pcoa[, !colnames(mtb_pcoa) %in% "Study.Group"]
cc_mtb$Study.Group <- as.factor(cc_mtb$Study.Group)
long_data <- cc_mtb %>%
  pivot_longer(cols = -Study.Group, names_to = "Compound", values_to = "Value")

# Perform Kruskal-Wallis test for each compound
compounds <- unique(long_data$Compound)
kruskal_results <- lapply(compounds, function(compound) {
  subset_data <- long_data %>% filter(Compound == compound)
  
  if (length(unique(subset_data$Study.Group)) < 2) {
    warning(paste("Skipping compound", compound, "due to insufficient group data"))
    return(NULL)
  }
  
  result <- kruskal.test(Value ~ Study.Group, data = subset_data)
  return(list(Compound = compound, Kruskal_Result = result))
})

results_mtb <- data.frame(
  Compound = character(),
  H_Statistic = numeric(),
  Degrees_Freedom = integer(),
  P_Value = numeric(),
  P_Value_FDR = numeric(),  
  Reject_Null = logical(),  
  stringsAsFactors = FALSE
)

for (res in kruskal_results) {
  if (!is.null(res)) {
    p_val <- res$Kruskal_Result$p.value
    p_val_fdr <- p.adjust(p_val, method = "fdr") # FDR corrected p-value
    reject_null <- p_val_fdr < 0.05
    
    results_mtb <- rbind(results_mtb, data.frame(
      Compound = res$Compound,
      H_Statistic = res$Kruskal_Result$statistic,
      Degrees_Freedom = res$Kruskal_Result$parameter,
      P_Value = p_val,
      P_Value_FDR = p_val_fdr,
      Reject_Null = reject_null
    ))
  }
}

num_rejected <- sum(results_mtb$Reject_Null)
num_accepted <- nrow(results_mtb) - num_rejected

cat("Number of null hypotheses rejected:", num_rejected, "\n")
cat("Number of null hypotheses accepted:", num_accepted, "\n")
```

# Remove insignificant features
```{r}
non_significant_compounds <- results_mtb$Compound[results_mtb$Reject_Null == FALSE]
mtb_filtered <- mtb_pcoa[, !names(mtb_pcoa) %in% non_significant_compounds]
print(mtb_filtered)
#the whole process left us with 103 microbes for machine learning analysis
```

#Save the datasets
```{r}
#microbiome
write.csv(gene_filtered, "genera_processed_CC.tsv", row.names = FALSE, col.names = TRUE)
#metabolome
write.csv(mtb_filtered, "mtb_processed_CC.tsv",row.names = FALSE, col.names = TRUE)
```



