---
title: "IBD Preprocessing"
author: "Daryll Susan Philip"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/DaryllPhilip/Desktop/THESIS /IBD")
```

```{r}
getwd()
```

#Load required packages
```{r}
library(tidyverse)
library(dplyr)
library(vegan)
library(tidyr)
library(naniar)
library(ape)
library(ggplot2)
library(FactoMineR)
library(factoextra)
library(stats)
library(openxlsx)
```


#Load the datasets
```{r}
#IBD microbes
gene_ibd = read.csv("genera_IBD.tsv", header = TRUE, sep = "\t", stringsAsFactors = FALSE)
head(gene_ibd)

#IBD metabolites
mtb_ibd = read.xlsx("mtb.xlsx")
mtb_ibd

```


#Microbiome IBD


```{r}
# Count the number of patients in each group
patient_counts <- gene_ibd %>%
  group_by(Study.Group) %>%
  summarise(Count = n())

print(patient_counts)
ibd_count <- sum(gene_ibd$Study.Group == "IBD")
healthy_count <- sum(gene_ibd$Study.Group == "Healthy")


cat("Number of IBD patients:", ibd_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

```



#Check for missing values
```{r}

missing_summary <- gene_ibd %>%
  summarise_all(~ sum(is.na(.))) %>%
  gather(key = "Column", value = "MissingValues") %>%
  mutate(MissingPercentage = (MissingValues / nrow(gene_ibd)) * 100)
missing_summary <- naniar::miss_var_summary(gene_ibd)

print(missing_summary)
naniar::gg_miss_var(gene_ibd)

#there were no missing values
```


#Identifying sparse features

```{r}

sparsity <- colSums(gene_ibd == 0) / nrow(gene_ibd) * 100
bins <- cut(sparsity, breaks = seq(0, 110, by = 10), include.lowest = TRUE, right = FALSE, labels = FALSE)
sparsity_counts <- table(bins)
bin_labels <- c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")
total_features <- length(sparsity)
surviving_features <- sum(bins == 1)  # bins == 1 corresponds to the "0%" bin

# Print the table
print("Sparsity (%) | Number of Features")
for (i in seq_along(bin_labels)) {
  count <- ifelse(i %in% names(sparsity_counts), sparsity_counts[as.character(i)], 0)
  cat(bin_labels[i], "|", count, "\n")
}

cat("\nTotal number of microbiome features:", total_features, "\n")
cat("Number of surviving microbiome features (non-sparse):", surviving_features, "\n")
```

#Remove the sparse microbiome features

```{r}
sparsity <- colSums(gene_ibd == 0) / nrow(gene_ibd) * 100
gene_new <- gene_ibd[, sparsity <= 40]
sparsity_filtered <- colSums(gene_new == 0) / nrow(gene_new) * 100
bins_filtered <- cut(sparsity_filtered, breaks = seq(0, 110, by = 10), include.lowest = TRUE, right = FALSE, labels = FALSE)
sparsity_counts_filtered <- table(bins_filtered)
bin_labels <- c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")

total_features_filtered <- length(sparsity_filtered)
surviving_features_filtered <- sum(bins_filtered == 1)
print("Sparsity (%) | Number of Features")
for (i in seq_along(bin_labels)) {
  count <- ifelse(i %in% names(sparsity_counts_filtered), sparsity_counts_filtered[as.character(i)], 0)
  cat(bin_labels[i], "|", count, "\n")
}

cat("\nTotal number of microbiome features after filtering:", total_features_filtered, "\n")
gene_new
```

#min-max scale the features

```{r}
min_max_scaling <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

gene_new <- gene_new %>%
  mutate(across(where(is.numeric), min_max_scaling))
gene_new
```


#PCoA to visualise the outliers (microbiome)

```{r}

gene_new <- gene_new[, !(names(gene_new) %in% c("Study.Group"))]
bray_curtis_dist <- vegdist(gene_new, method = "bray")
pcoa_results <- pcoa(bray_curtis_dist)
coordinates <- pcoa_results$vectors[, 1:2]
colnames(coordinates) <- c("Axis.1", "Axis.2")
pcoa_df <- as.data.frame(coordinates)
pcoa_df$Study.Group <- gene_new$Study.Group  
pcoa_df$Study.Group <- as.factor(pcoa_df$Study.Group)


pcoa_df <- pcoa_df %>%
  mutate(OutlierStatus = ifelse(Outlier, "Outlier", "Non-Outlier"))
print(table(pcoa_df$OutlierStatus))

#Plot PCoA with outliers highlighted
ggplot(pcoa_df, aes(x = Axis.1, y = Axis.2, color = Study.Group)) +
  geom_point(aes(shape = OutlierStatus), size = 3) +  # Shape based on outlier status
  stat_ellipse(aes(group = Study.Group, color = Study.Group), type = "t", level = 0.95) +
  scale_color_manual(values = c("IBD" = "#D95F02", "Healthy" = "#1B9E77")) +
  labs(title = "PCoA For Genera (Bray-Curtis) with Outliers",
       x = paste("PCoA1 (", round(100 * pcoa_results$values$Relative_eig[1], 2), "%)", sep = ""),
       y = paste("PCoA2 (", round(100 * pcoa_results$values$Relative_eig[2], 2), "%)", sep = ""),
       color = "Group", shape = "Outlier Status") +
  theme_minimal()

```

#Remove the outliers and replot the graph 
```{r}

non_outlier_indices <- which(pcoa_df$OutlierStatus == "Non-Outlier")
cleaned_gene <- gene_new[non_outlier_indices, ]
gen_cleaned <- cleaned_gene[, !(names(cleaned_gene) %in% c("Study.Group"))]
bray_curtis_dist_cleaned <- vegdist(gen_cleaned, method = "bray")

# Perform PCoA on the cleaned Bray-Curtis dissimilarity matrix
pcoa_results_cleaned <- pcoa(bray_curtis_dist_cleaned)
coordinates_cleaned <- pcoa_results_cleaned$vectors[, 1:2]
colnames(coordinates_cleaned) <- c("Axis.1", "Axis.2")
pcoa_df_cleaned <- as.data.frame(coordinates_cleaned)
pcoa_df_cleaned$Study.Group <- cleaned_gene$Study.Group  # Ensure this column exists in 'cleaned_gene'
pcoa_df_cleaned$Study.Group <- as.factor(pcoa_df_cleaned$Study.Group)

#Plot PCoA for cleaned dataset with outliers removed
ggplot(pcoa_df_cleaned, aes(x = Axis.1, y = Axis.2, color = Study.Group)) +
  geom_point(size = 3) +  
  stat_ellipse(aes(group = Study.Group, color = Study.Group), type = "t", level = 0.95) +
  scale_color_manual(values = c("IBD" = "#D95F02", "Healthy" = "#1B9E77")) +
  labs(title = "PCoA For IBD Genera (Bray-Curtis)",
       x = paste("PCoA1 (", round(100 * pcoa_results_cleaned$values$Relative_eig[1], 2), "%)", sep = ""),
       y = paste("PCoA2 (", round(100 * pcoa_results_cleaned$values$Relative_eig[2], 2), "%)", sep = ""),
       color = "Group") +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.x = element_text(size = 14),  
    axis.title.y = element_text(size = 14), 
    legend.title = element_text(size = 12)  
  )

```

#Check for number of ibd and healthy patients for microbiome dataset
```{r}
gene_pcoa = cleaned_gene
patient_counts <- gene_pcoa %>%
  group_by(Study.Group) %>%
  summarise(Count = n())
print(patient_counts)


ibd_count <- sum(gene_pcoa$Study.Group == "IBD")
healthy_count <- sum(gene_pcoa$Study.Group == "Healthy")

cat("Number of IBD patients:", ibd_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

# 211 Samples from the microbiome dataset were considered for final analysis from the previous 220
```

#Mann Whitney U test/ Wilcoxon Sum Rank Test For Genera/Microbiome

```{r}

long_data <- gene_pcoa %>%
  pivot_longer(cols = -c(Study.Group), names_to = "Compound", values_to = "Value")
p_values <- list()
compounds <- unique(long_data$Compound)
p_values_vec <- unlist(p_values)

# Apply FDR correction using Benjamini-Hochberg procedure
p_adjusted <- p.adjust(p_values_vec, method = "fdr")

# Create a result data frame
results <- data.frame(
  Compound = names(p_values),
  Original_P_Value = p_values_vec,
  FDR_Corrected_P_Value = p_adjusted,
  Reject_Null = p_adjusted < 0.05
)

print(results)
count_reject_null <- table(results$Reject_Null)
print(count_reject_null)


num_rejected <- count_reject_null['TRUE']
num_accepted <- count_reject_null['FALSE']

cat("Number of null hypotheses rejected (significant features):", num_rejected, "\n")
cat("Number of null hypotheses accepted (non-significant features):", num_accepted, "\n")

```

#Remove insignificant features
```{r}

significant_results <- results %>% filter(Reject_Null == TRUE)
print(significant_results)
gene_filtered <- gene_pcoa %>%
  select(Study.Group,  all_of(significant_results$Compound))


print(gene_filtered)
dim(gene_filtered)

#the whole process left us with 1089 microbes for machine learning analysis
```


#Metabolites(IBD)


```{r}
mtb_ibd <- mtb_ibd[, !grepl(":.NA$", colnames(mtb_ibd))]
mtb_ibd
```



```{r}
# Count the number of patients in each group
patient_counts <- mtb_ibd %>%
  group_by(Study.Group) %>%
  summarise(Count = n())

print(patient_counts)
ibd_count <- sum(mtb_ibd$Study.Group == "IBD")
healthy_count <- sum(mtb_ibd$Study.Group == "Healthy")


cat("Number of IBD patients:", ibd_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

```



#Check for missing values
```{r}

missing_summary <- mtb_ibd %>%
  summarise_all(~ sum(is.na(.))) %>%
  gather(key = "Column", value = "MissingValues") %>%
  mutate(MissingPercentage = (MissingValues / nrow(mtb_ibd)) * 100)
missing_summary <- naniar::miss_var_summary(mtb_ibd)

print(missing_summary)
naniar::gg_miss_var(mtb_ibd)

#there were no missing values
```


#Identifying sparse features
```{r}

sparsity <- colSums(mtb_ibd == 0) / nrow(mtb_ibd) * 100
bins <- cut(sparsity, breaks = seq(0, 110, by = 10), include.lowest = TRUE, right = FALSE, labels = FALSE)
sparsity_counts <- table(bins)
bin_labels <- c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")
total_features <- length(sparsity)
surviving_features <- sum(bins == 1)  # bins == 1 corresponds to the "0%" bin

# Print the table
print("Sparsity (%) | Number of Features")
for (i in seq_along(bin_labels)) {
  count <- ifelse(i %in% names(sparsity_counts), sparsity_counts[as.character(i)], 0)
  cat(bin_labels[i], "|", count, "\n")
}

cat("\nTotal number of metabolome features:", total_features, "\n")
cat("Number of surviving metabolome features (non-sparse):", surviving_features, "\n")
```

#Remove the sparse metabolome features
```{r}
sparsity <- colSums(mtb_ibd == 0) / nrow(mtb_ibd) * 100
mtb_new <- mtb_ibd[, sparsity <= 80]
sparsity_filtered <- colSums(mtb_new == 0) / nrow(mtb_new) * 100
bins_filtered <- cut(sparsity_filtered, breaks = seq(0, 110, by = 10), include.lowest = TRUE, right = FALSE, labels = FALSE)
sparsity_counts_filtered <- table(bins_filtered)
bin_labels <- c("0%", "10%", "20%", "30%", "40%", "50%", "60%", "70%", "80%", "90%", "100%")

total_features_filtered <- length(sparsity_filtered)
surviving_features_filtered <- sum(bins_filtered == 1)
print("Sparsity (%) | Number of Features")
for (i in seq_along(bin_labels)) {
  count <- ifelse(i %in% names(sparsity_counts_filtered), sparsity_counts_filtered[as.character(i)], 0)
  cat(bin_labels[i], "|", count, "\n")
}

cat("\nTotal number of metabolome features after filtering:", total_features_filtered, "\n")
mtb_new
```

#Min-max scale the features
```{r}
min_max_scaling <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

mtb_new <- mtb_new %>%
  mutate(across(where(is.numeric), min_max_scaling))
mtb_new
```


#PCA for metabolites

```{r}


mtbs <- mtb_new[, !(names(mtb_new) %in% c("Study.Group"))]


mtbs <- mtbs[, apply(mtbs, 2, function(col) sd(col) != 0)]
pca_result_mtb <- prcomp(mtbs, scale. = TRUE)
fviz_eig(pca_result_mtb, addlabels = TRUE, ylim = c(0, 70))
explained_variance <- summary(pca_result_mtb)$importance[2, ]
pca_df <- as.data.frame(pca_result_mtb$x)



#Calculate Mahalanobis distance
mean_pca <- colMeans(pca_df[, 1:2])
cov_pca <- cov(pca_df[, 1:2])
mahalanobis_dist <- apply(pca_df[, 1:2], 1, function(x) mahalanobis(x, mean_pca, cov_pca))
cutoff <- qchisq(0.95, df = 2)
pca_df$Outlier <- mahalanobis_dist > cutoff


pca_df <- pca_df %>%
  mutate(OutlierStatus = ifelse(Outlier, "Outlier", "Non-Outlier"))

#Plot PCA 
ggplot(pca_df, aes(x = PC1, y = PC2, color = Study.Group, shape = OutlierStatus)) +
  geom_point(size = 3) +
  scale_shape_manual(values = c("Non-Outlier" = 16, "Outlier" = 4)) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "PCA For Metabolites",
       x = paste("PC1 (", round(100 * explained_variance[1], 2), "%)", sep = ""),
       y = paste("PC2 (", round(100 * explained_variance[2], 2), "%)", sep = ""),
       color = "Study.Group",
       shape = "Outlier Status") +
  theme_minimal()

table(pca_df$OutlierStatus)

```

#Remove outliers and replot the graph

```{r}

pca_df_cleaned <- pca_df %>%
  filter(OutlierStatus == "Non-Outlier")

mtb_ibd <- mtb_new %>%
  filter(row.names(mtb_new) %in% row.names(pca_df_cleaned))

#Plot PCA with outliers removed
ggplot(pca_df_cleaned, aes(x = PC1, y = PC2, color = Study.Group)) +
  geom_point(size = 3) +
  scale_color_brewer(palette = "Dark2") +
  labs(title = "PCA For IBD Metabolites ",
       x = paste("PC1 (", round(100 * explained_variance[1], 2), "%)", sep = ""),
       y = paste("PC2 (", round(100 * explained_variance[2], 2), "%)", sep = ""),
       color = "Group") +
  theme_minimal() +
theme(
    plot.title = element_text(size = 16),  
    axis.title.x = element_text(size = 14), 
    axis.title.y = element_text(size = 14), 
    legend.title = element_text(size = 12)  
  )
```

#Check for number of IBD and Healthy patients for metabolome dataset
```{r}
mtb_pcoa = mtb_ibd
patient_counts <- mtb_pcoa %>%
  group_by(Study.Group) %>%
  summarise(Count = n())
print(patient_counts)


ibd_count <- sum(mtb_pcoa$Study.Group == "IBD")
healthy_count <- sum(mtb_pcoa$Study.Group == "Healthy")

cat("Number of IBD patients:", ibd_count, "\n")
cat("Number of Healthy patients:", healthy_count, "\n")

# 213 samples from the metabolome dataset were considered for final analysis from the previous 220
```




#Mann Whitney U test/ Wilcoxon Sum Rank Test For Metabolome 

```{r}

long_data <- mtb_pcoa %>%
  pivot_longer(cols = -c(Study.Group), names_to = "Compound", values_to = "Value")
p_values <- list()
compounds <- unique(long_data$Compound)
p_values_vec <- unlist(p_values)

# Apply FDR correction using Benjamini-Hochberg procedure
p_adjusted <- p.adjust(p_values_vec, method = "fdr")

# Create a result data frame
results <- data.frame(
  Compound = names(p_values),
  Original_P_Value = p_values_vec,
  FDR_Corrected_P_Value = p_adjusted,
  Reject_Null = p_adjusted < 0.05
)

print(results)
count_reject_null <- table(results$Reject_Null)
print(count_reject_null)


num_rejected <- count_reject_null['TRUE']
num_accepted <- count_reject_null['FALSE']

cat("Number of null hypotheses rejected (significant features):", num_rejected, "\n")
cat("Number of null hypotheses accepted (non-significant features):", num_accepted, "\n")

```

#Remove insignificant features
```{r}

significant_results <- results %>% filter(Reject_Null == TRUE)
print(significant_results)
mtb_filtered <- mtb_pcoa %>%
  select(Study.Group,  all_of(significant_results$Compound))


print(mtb_filtered)
dim(mtb_filtered)

#the whole process left us with 259 metabolites for machine learning analysis
```



#Save the datasets
```{r}
#microbiome
write.csv(gene_filtered, "genera_processed_IBD.tsv", row.names = FALSE, col.names = TRUE)
#metabolome
write.csv(mtb_filtered, "mtb_processed_IBD.tsv",row.names = FALSE, col.names = TRUE)
```