---
title: "Extra visualizations"
author: "Daryll Susan Philip"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Load required packages
```{r}
library(tidyverse)      
library(forcats)         
library(glue) 
library(devtools)
library(ggplot2)
```

#Gastric Cancer Model

```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accura
               cy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.92, 0.79, 0.81, 0.91, 0.80, 0.69,  # XGBoost metrics
            0.94, 0.79, 0.71, 0.91, 0.80, 0.69,  # Random Forest metrics
            0.86, 0.83, 0.90, 0.82, 0.86, 0.92)  # LASSO metrics
)

# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  # Offset each model's group by 6 IDs
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO")) 

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "top",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For Main Gastric Cancer Model") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))

```


#GC Prediction on IBD

```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.73, 0.76, 0.74, 0.76, 0.69, 0.41,  # XGBoost metrics
            0.77, 0.76, 0.78, 0.76, 0.68, 0.09,  # Random Forest metrics
            0.75, 0.76, 0.74, 0.76, 0.74, 0.34)  # LASSO metrics
)
# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For Gastric Cancer Features on IBD") 
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```


#GC Predictions on CRC

```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.59, 0.59, 0.60, 0.59, 0.59, 0.08,  # XGBoost metrics
            0.63, 0.61, 0.66, 0.61, 0.61, 0.93,  # Random Forest metrics
            0.58, 0.56, 0.56, 0.56, 0.53, 0.79)  # LASSO metrics
)
# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 19),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For Gastric Cancer Features for CC") +
  scale_y_continuous(limits = c(-10, 100)) + 
  
  # Guide for legend with a clearer layout
  guides(fill = guide_legend(title = "Models"))
```

#Colon Cancer Model

```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.72, 0.67, 0.67, 0.67, 0.67, 0.64,  # XGBoost metrics
            0.75, 0.66, 0.65, 0.66, 0.66, 0.72,  # Random Forest metrics
            0.70, 0.67, 0.67, 0.67, 0.54, 0.69)  # LASSO metrics
)
# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For Colon Cancer Main Model") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```


#CRC predictions on GC

```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.85, 0.69, 0.73, 0.69, 0.68, 0.86,  # XGBoost metrics
            0.86, 0.76, 0.79, 0.76, 0.76, 0.88,  # Random Forest metrics
            0.70, 0.69, 0.71, 0.69, 0.69, 0.62)  # LASSO metrics
)
# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  # Offset each model's group by 6 IDs
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  # Reorder levels

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For Colon Cancer features in GC") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```


#CRC predictions on IBD

```{r}

data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.58, 0.75, 0.68, 0.75, 0.64, 0.61,  # XGBoost metrics
            0.52, 0.75, 0.81, 0.75, 0.64, 0.93,  # Random Forest metrics
            0.65, 0.75, 0.81, 0.75, 0.64, 0.86)  # LASSO metrics
)
# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  # Offset each model's group by 6 IDs
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  # Reorder levels

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For Colon Cancer features in IBD") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```

#IBD model
```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.87, 0.82, 0.82, 0.82, 0.82, 0.64,  # XGBoost metrics
            0.93, 0.84, 0.83, 0.84, 0.83, 0.64,  # Random Forest metrics
            0.84, 0.78, 0.80, 0.78, 0.78, 0.64)  # LASSO metrics
)

# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  # Offset each model's group by 6 IDs
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  # Reorder levels

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 6, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For IBD Main Model") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```

#IBD predictions on GC

```{r}
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.50, 0.56, 0.75, 0.56, 0.56, 0,  # XGBoost metrics
            0.66, 0.56, 0.75, 0.56, 0.56, 0,  # Random Forest metrics
            0.64, 0.56, 0.56, 0.56, 0.56, 0.05)  # LASSO metrics
)
# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  # Offset each model's group by 6 IDs
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  # Reorder levels

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 5.5, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For IBD features for GC") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```

#IBD predictions on CRC

```{r}

data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.53, 0.54, 0.75, 0.54, 0.54, 1,  # XGBoost metrics
            0.53, 0.54, 0.75, 0.54, 0.54, 1,  # Random Forest metrics
            0.57, 0.57, 0.57, 0.57, 0.57, 0.01)  # LASSO metrics
)

# Adding IDs to separate sections for each model
data <- data %>%
  mutate(id = row_number() + rep(0:2 * 6, each = 6)) %>%  # Offset each model's group by 6 IDs
  mutate(group = fct_relevel(group, "XGBoost", "Random Forest", "LASSO"))  # Reorder levels

# Define custom colors for each model
custom_colors <- c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")

# Plot
ggplot(data, aes(x = factor(id), y = value * 100, fill = group)) +
  geom_bar(aes(y = 100), stat = "identity", width = 1, color = "white", alpha = 0.5) +
  geom_bar(stat = "identity", width = 1, color = "white", size = 1.1) +
  geom_label(aes(label = round(value * 100, 2), fill = group), size = 5.7, 
             color = "white", show.legend = FALSE) +
  
  # Circular layout
  coord_polar(start = 0) +
  scale_fill_manual(values = custom_colors) +
  
  # Styling adjustments
  theme_void() +
  theme(
    legend.position = "right",
    legend.title = element_blank(),
    legend.text = element_text(size = 18),
    plot.title = element_text(hjust = 0.5, size = 19, face = "bold", color = "#333333"),
    plot.margin = margin(5, 5, 5, 5)
  ) +
  
  # Plot title
  labs(title = "Model Metrics For IBD features for CC") +
  scale_y_continuous(limits = c(-10, 100)) + 
  guides(fill = guide_legend(title = "Models"))
```



#GASTRIC CANCER VALIDATION BAR GRAPH 

#Microbiome

```{r, fig.width=6, fig.height=4}

# Create the data frame
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.58, 0.80, 0.84, 0.80, 0.71, 0,  # XGBoost metrics
            0.88, 0.90, 0.90, 0.90, 0.90, 0,  # Random Forest metrics
            0.50, 0.80, 0.84, 0.80, 0.71, 1)  # LASSO metrics
)
# Reorder the 'name' column to ensure the desired order of metrics
data$name <- factor(data$name, levels = c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"))
data$group <- factor(data$group, levels = c("XGBoost", "Random Forest", "LASSO"))

# Plot
ggplot(data, aes(x = name, y = value, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "",
       fill = "Model") +
  scale_fill_manual(values = c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 0.5, size = 18, colour= 'black'),
    axis.text.y = element_text(size = 18, colour= 'black'),
    legend.title = element_text(size = 18, face = "bold"),   
    legend.text = element_text(size = 18),
    legend.position = "bottom",  
    legend.direction = "horizontal"
  )

```



#Metabolite

```{r, fig.width=6, fig.height=4}

# Create the data frame
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.50, 0.95, 0.95, 0.95, 0.92, 1,  # XGBoost metrics
            0.50, 0.95, 0.95, 0.95, 0.92, 1,  # Random Forest metrics
            0.50, 0.95, 0.95, 0.95, 0.92, 1)  # LASSO metrics
)

# Reorder the 'name' column to ensure the desired order of metrics
data$name <- factor(data$name, levels = c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"))
data$group <- factor(data$group, levels = c("XGBoost", "Random Forest", "LASSO"))

# Plot
ggplot(data, aes(x = name, y = value, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "",
       fill = "Model") +
  scale_fill_manual(values = c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 0.5, size = 18, colour= 'black'),
    axis.text.y = element_text(size = 18, colour= 'black'),
    legend.title = element_text(size = 18, face = "bold"), 
    legend.text = element_text(size = 18)
  )

```



#Colon Cancer

#Microbiome

```{r}

# Create the data frame
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.51, 0.74, 0.81, 0.74, 0.74, 0,  # XGBoost metrics
            0.50, 0.74, 0.81, 0.74, 0.74, 0,  # Random Forest metrics
            0.50, 0.74, 0.74, 0.74, 0.74, 0)  # LASSO metrics
  
)

# Reorder the 'name' column to ensure the desired order of metrics
data$name <- factor(data$name, levels = c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"))
data$group <- factor(data$group, levels = c("XGBoost", "Random Forest", "LASSO"))

# Plot
ggplot(data, aes(x = name, y = value, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "",
       fill = "Model") +
  scale_fill_manual(values = c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")) + 
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 0.5, size = 18, colour= 'black'),
    axis.text.y = element_text(size = 18, colour= 'black'),
    legend.title = element_text(size = 18, face = "bold"),   
    legend.text = element_text(size = 18),
    legend.position = "bottom",  
    legend.direction = "horizontal"
  )
```


#Metabolite

```{r, fig.width=6, fig.height=4}

# Create the data frame
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.46, 0.74, 0.81, 0.74, 0.74, 0.98,  # XGBoost metrics
            0.50, 0.74, 0.81, 0.74, 0.63, 0.75,  # Random Forest metrics
            0.53, 0.74, 0.69, 0.74, 0.68, 0.92)  # LASSO metrics
)

# Reorder the 'name' column to ensure the desired order of metrics
data$name <- factor(data$name, levels = c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"))
data$group <- factor(data$group, levels = c("XGBoost", "Random Forest", "LASSO"))

# Plot
ggplot(data, aes(x = name, y = value, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "",
       fill = "Model") +
  scale_fill_manual(values = c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 0.5, size = 18, colour= 'black'),
    axis.text.y = element_text(size = 18, colour= 'black'),
    legend.title = element_text(size = 18, face = "bold"),   
    legend.text = element_text(size = 18),
    legend.position = "bottom",
    legend.direction = "horizontal"
  )
```

#IBD

#Microbiome

```{r}
# Create the data frame
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.59, 0.75, 0.76, 0.75, 0.75, 0.09,  # XGBoost metrics
            0.60, 0.74, 0.70, 0.74, 0.74, 0.18,  # Random Forest metrics
            0.56, 0.73, 0.80, 0.73, 0.73, 0.02)  # LASSO metrics
)

# Reorder the 'name' column to ensure the desired order of metrics
data$name <- factor(data$name, levels = c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"))
data$group <- factor(data$group, levels = c("XGBoost", "Random Forest", "LASSO"))

# Plot
ggplot(data, aes(x = name, y = value, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "",
       fill = "Model") +
  scale_fill_manual(values = c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 0.5, size = 18, colour= 'black'),
    axis.text.y = element_text(size = 18, colour= 'black'),
    legend.title = element_text(size = 18, face = "bold"),   
    legend.text = element_text(size = 18),
    legend.position = "bottom", 
    legend.direction = "horizontal"
  )

```


#Metabolite

```{r, fig.width=6, fig.height=4}

# Create the data frame
data <- data.frame(
  group = c(rep("XGBoost", 6), rep("Random Forest", 6), rep("LASSO", 6)),
  name = rep(c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"), 3),
  value = c(0.47, 0.74, 0.81, 0.74, 0.74, 0.11,  # XGBoost metrics
            0.47, 0.74, 0.81, 0.74, 0.63, 0.97,  # Random Forest metrics
            0.50, 0.75, 0.81, 0.75, 0.64, 0.14)  # LASSO metrics
)

# Reorder the 'name' column to ensure the desired order of metrics
data$name <- factor(data$name, levels = c("ROC AUC", "Accuracy", "Precision", "Recall", "F1", "Specificity"))
data$group <- factor(data$group, levels = c("XGBoost", "Random Forest", "LASSO"))

# Plot
ggplot(data, aes(x = name, y = value, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "red") +
  labs(title = "",
       fill = "Model") +
  scale_fill_manual(values = c("XGBoost" = "#1A78CF", "Random Forest" = "#00A651", "LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(hjust = 0.5, size = 18, colour= 'black'),
    axis.text.y = element_text(size = 18, colour= 'black'),
    legend.title = element_text(size = 18, face = "bold"),   
    legend.text = element_text(size = 18),
    legend.position = "bottom",  
    legend.direction = "horizontal"
  )
```




#Top Features
#GC microbiome

```{r}

# Create the data frame
data <- data.frame(
  name = factor(c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30"), 
                levels = c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30")),  
  value = c(0.74, 0.9, 0.97, 0.96, 0.96, 0.94)   
)

# Plot
ggplot(data, aes(y = name, x = value, fill = "Random Forest")) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "red") +
  labs(
    title = "",
    x = "Performance Metric",  
    y = "Top Features",        
    fill = "Model"
  ) +
  scale_fill_manual(values = c("Random Forest" = "#00A651")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.text.y = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 18),
    legend.position = "bottom",  # Position legend at the bottom
    legend.direction = "horizontal"
  )
```

#GC metabolome

```{r}

# Create the data frame
data <- data.frame(
  name = factor(c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30"), 
                levels = c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30")),  
  value = c(0.8, 0.89, 0.86, 0.85, 0.83, 0.98)  # 
)

# Plot
ggplot(data, aes(y = name, x = value, fill = "LASSO")) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "",
    x = "Performance Metric",  
    y = "Top Features",        
    fill = "Model"
  ) +
  scale_fill_manual(values = c("LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.text.y = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 18),
    legend.position = "bottom",  
    legend.direction = "horizontal"
  )

```


#CRC Microbes 

```{r}

# Create the data frame
data <- data.frame(
  name = factor(c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30"), 
                levels = c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30")),
  value = c(0.53, 0.51, 0.54, 0.53, 0.53, 0.51)  # 
)

# Plot
ggplot(data, aes(y = name, x = value, fill = "Random Forest")) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "",
    x = "Performance Metric",  # Set x-axis label
    y = "Top Features",        # Set y-axis label
    fill = "Model"
  ) +
  scale_fill_manual(values = c("Random Forest" = "#00A651")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.text.y = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 18),
    legend.position = "bottom",  
    legend.direction = "horizontal"
  )

```



#CRC Metabolome

```{r}
# Create the data frame
data <- data.frame(
  name = factor(c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30"), 
                levels = c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30")),  
  value = c(0.55, 0.59, 0.60, 0.61, 0.61, 0.61)  # 
)

# Plot
ggplot(data, aes(y = name, x = value, fill = "LASSO")) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "",
    x = "Performance Metric", 
    y = "Top Features",        
    fill = "Model"
  ) +
  scale_fill_manual(values = c("LASSO" = "#D70232")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.text.y = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 18),
    legend.position = "bottom", 
    legend.direction = "horizontal"
  )
```


#IBD microbiome

```{r}

# Create the data frame
data <- data.frame(
  name = factor(c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30"), 
                levels = c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30")),  
  value = c(0.88, 0.87, 0.89, 0.85, 0.84, 0.85)  # 
)

# Plot
ggplot(data, aes(y = name, x = value, fill = "Random Forest")) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "",
    x = "Performance Metric",  
    y = "Top Features",      
    fill = "Model"
  ) +
  scale_fill_manual(values = c("Random Forest" = "#00A651")) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.text.y = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 18),
    legend.position = "bottom",  
    legend.direction = "horizontal"
  )

```

#IBD metabolome

```{r}

# Create the data frame
data <- data.frame(
  name = factor(c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30"), 
                levels = c("Top 5", "Top 10", "Top 15", "Top 20", "Top 25", "Top 30")),  
  value = c(0.84, 0.87, 0.87, 0.88, 0.87, 0.88)  # 
)

# Plot
ggplot(data, aes(y = name, x = value, fill = "Random Forest")) +
  geom_bar(stat = "identity", position = position_dodge(), color = "black") +
  geom_vline(xintercept = 0.5, linetype = "dashed", color = "black") +
  labs(
    title = "",
    x = "Performance Metric",
    y = "Top Features",        
    fill = "Model"
  ) +
  scale_fill_manual(values = c("Random Forest" = "#00A651")) +  
  theme_minimal() +
  theme(
    axis.text.x = element_text(size = 18, colour = "black"),
    axis.text.y = element_text(size = 18, colour = "black"),
    axis.title.x = element_text(size = 18, face = "bold"),
    axis.title.y = element_text(size = 18, face = "bold"),
    legend.title = element_text(size = 18, face = "bold"),
    legend.text = element_text(size = 18),
    legend.position = "bottom", 
    legend.direction = "horizontal"
  )

```




